25 Years of Water Quality Change in Rhode Island Lakes and Ponds

### 2018-12-19 manuscript/manuscript.Rmd
figures added to paper

The study area for this analyis includes data in lakes and ponds in the state of Rhode Island that were collected by the University of Rhode Island's Watershed Watch (URIWW) program (Figure X).  MORE DETAILS ON RIWW.

In particular we selected data from the RIWW data that matched the following criteria: 1) were sampled between 1990 and 2016, 2) were sampled in May to October, 3) and were sampled at a depth of 2 meters or less. For this analysis we were interested in trends in lake temperature, total nitorgen, total phosphorus, and chlorophyll *a* and for each of these parameters we further filtered the data to select sites that had at least 10 years of data.  The final dataset used in our analysis included `r nrow(ww_avail_data_locations)` lakes and ponds which had approximately `r length(temp_10yr_sites)` samples for temperature, `r length(chla_10yr_sites)` samples for chlorophyll *a*, `r length(total_n_10yr_sites)` samples for Total Nitrogen (TN), and `r length(total_p_10yr_sites)` samples for Total Phosphorus (TP).

@@

[TN, TP, CHLA, TEMP Figures]

@@

>>>>

The study area for this analyis includes data in lakes and ponds in the state of Rhode Island that were collected by the University of Rhode Island's Watershed Watch (URIWW) program (`r fig_num("ww_map", display = "cite", caption = "Map of URI Watershed Watch lake and pond sampling sites")`).  MORE DETAILS ON RIWW.

![`r fig_num("ww_map", caption = "Map of URI Watershed Watch lake and pond sampling sites")`](../figures/ww_map.jpg)

In particular we selected data from the RIWW data that matched the following criteria: 1) were sampled between 1990 and 2016, 2) were sampled in May to October, 3) and were sampled at a depth of 2 meters or less. For this analysis we were interested in trends in lake temperature, total nitorgen, total phosphorus, and chlorophyll *a* and for each of these parameters we further filtered the data to select sites that had at least 10 years of data for a given parameter.  The final dataset used in our analysis included `r nrow(ww_avail_data_locations)` lakes and ponds which had approximately `r length(temp_10yr_sites)` samples for temperature, `r length(chla_10yr_sites)` samples for chlorophyll *a*, `r length(total_n_10yr_sites)` samples for Total Nitrogen (TN), and `r length(total_p_10yr_sites)` samples for Total Phosphorus (TP).  Of the `r nrow(ww_avail_data_locations)` sampling sites, `r sum(avail_data$params_avail == "chla, total_p, total_n, temp")` had data for all 4 parameters.

@@

![`r fig_num(name = "temp_ri", caption = "Yearly trend over 20+ years of lake temperature in Rhode Island lakes and ponds.")`](../figures/ww_temp_trends.jpg)

![`r fig_num(name = "tn_ri", caption = "Yearly trend over 20+ years of Total Nitrogen (average z-score) in Rhode Island lakes and ponds.")`](../figures/ww_tn_trends.jpg)

![`r fig_num(name = "tp_ri", caption = "Yearly trend over 20+ years of Total Phosphorus (average z-score) in Rhode Island lakes and ponds.")`](../figures/ww_tp_trends.jpg)

![`r fig_num(name = "chla_ri", caption = "Yearly trend over 20+ years of Chlorphyll *a* (average z-score) in Rhode Island lakes and ponds.")`](../figures/ww_chla_trends.jpg)

@@

![`r fig_num(name = "tn_lagos", caption = "Yearly trend over 20+ years of Total Nitrogen (average z-score) in the LAGOSNE states.")`](../figures/lagos_tn_trends.jpg)

![`r fig_num(name = "tp_lagos", caption = "Yearly trend over 20+ years of Total Phosphorus (average z-score) in the LAGOSNE states.")`](../figures/lagos_tp_trends.jpg)

![`r fig_num(name = "chla_lagos", caption = "Yearly trend over 20+ years of Chlorophyll *a* (average z-score) in the LAGOSNE states.")`](../figures/lagos_chla_trends.jpg)
